Choosing and rating harmonies

### 2020-08-14 main.Rmd
paper structure

opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = "figure/", fig.align = "center", fig.show = "hold",
  cache = TRUE, cache.path = "cache/",
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
read_chunk('check_gt_mnylvls.R')
```

@@

# Idea

@@

# Computing distances

@@

Theoretical: JS distances are distributed as chi-squared with $m$ df where we discretize the continuous distribution with $m$ discrete values. Taking sample percentiles to approximate the integral would mean taking $m = 99$.
With large $m$, chi-squared is asymptotically normal by the CLT. Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution. Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function. $n$ is the number of pairwise comparisons being made.

Empirical: Distribution of JS distances is assumed to be normal but the mean and variance are estimated from the sample, rather than deducing it from the number of discretization used to approximate the continuous distribution.

# Choose thresholds for harmonies through permutation test

@@

# Does normalization work? -  Minimal examples

Consider two cyclic granularities $A$ and $B$ with $2$ and $3$categories. Thus, the harmony table consisting of all possible harmony pairs (assuming all pairs are harmonies), would look like the following:

```{r harmony_min}
# harmonies <- tibble::tibble(facet_variable = c("A", "B", "A", "C", "B", "C"),
#                             x_variable  = c("B","A", "C", "A", "C", "B"),
#                             facet_levels = c(2, 3, 2, 4, 3, 4),
#                             x_levels = c(3, 2, 4, 2, 4, 3))

harmonies <- tibble::tibble(facet_variable = c("A", "B"),
                            x_variable  = c("B","A"),
                            facet_levels = c(2, 3),
                            x_levels = c(3, 2))

harmonies %>% knitr::kable()
```

The output table has the value of MMPD (normalized median maximum pairwise distances), gt_MMPD(global threshold of MMPD indicator).

@@

*Conclusion*: The test with MMPD rejects the harmony pair even for which distribution of x-axis categories are significantly different. But the test with only maximum selects the pair with varying x-axis categories.

## Scenario 4: Cumulative 3 levels with 2, 7 and 11 and testing level and power

```{r samenull_3levels}

*Conclusion*: With 3 levels the test incorrectly chooses 1 harmony pair with similar distribution. The harmony pair which is displayed.

# Levels and Power of the test statistic

<!-- ```{r diffnull_3levels} -->

@@

*Conclusion*: The test with MMPD selects just one pair, as opposed to the test with maximum. This needs to be checked against what we expect from the test. The harmony pairs which are selected (either through MMPD or maximum) are displayed.

# Permutation test

@@

>>>>

opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = "figure/", fig.align = "center", fig.show = "hold",
  cache = TRUE, cache.path = "cache/",
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
read_chunk('main.R')
```

@@

```{r}
set.seed(321)
```

# Introduction

# Background

@@

# A new test statistic

## Computing distances

@@

### Theoretical evidence

JS distances are distributed as chi-squared with $m$ df where we discretize the continuous distribution with $m$ discrete values. Taking sample percentiles to approximate the integral would mean taking $m = 99$.
With large $m$, chi-squared is asymptotically normal by the CLT. Thus, by CLT, ${\chi^2}_{m} \tilde{} N(m, 2m)$, which would depend on the number of discretization used to approximate the continuous distribution. Then $b_n = 1-1/n$ quantile of the normal distribution and $a_n = 1/[n*\phi(b_n)]$ where $\phi$ is the normal density function. $n$ is the number of pairwise comparisons being made.

### Empirical evidence

Distribution of JS distances is assumed to be normal but the mean and variance are estimated from the sample, rather than deducing it from the number of discretization used to approximate the continuous distribution.

# The statistical test

## Threshold definition for test through permutation test

@@

## Level and power of the test

<!-- Consider two cyclic granularities $A$ and $B$ with $2$ and $3$categories. Thus, the harmony table consisting of all possible harmony pairs (assuming all pairs are harmonies), would look like the following: -->

```{r harmony_min}
# harmonies <- tibble::tibble(facet_variable = c("A", "B", "A", "C", "B", "C"),
#                             x_variable  = c("B","A", "C", "A", "C", "B"),
#                             facet_levels = c(2, 3, 2, 4, 3, 4),
#                             x_levels = c(3, 2, 4, 2, 4, 3))
# 
# harmonies <- tibble::tibble(facet_variable = c("A", "B"),
#                             x_variable  = c("B","A"),
#                             facet_levels = c(2, 3),
#                             x_levels = c(3, 2))
#                             
# harmonies %>% knitr::kable()
```

<!-- The output table has the value of MMPD (normalized median maximum pairwise distances), gt_MMPD(global threshold of MMPD indicator). -->

@@

*Conclusion*: The test indicates that both harmony pairs do not have significant variation.

```{r diffnull_7by11normal}
```

*Conclusion*: The test indicates that only the first harmony pair has significant variation.

```{r diffnull_7by11normal2}
```

*Conclusion*: The test indicates that only the first harmony pair has significant variation.

<!-- ## Scenario 4: Cumulative 3 levels with 2, 7 and 11 and testing level and power -->

<!-- ```{r samenull_3levels} -->

@@

<!-- *Conclusion*: With 3 levels the test incorrectly chooses 1 harmony pair with similar distribution. The harmony pair which is displayed. -->

<!-- <!-- ```{r diffnull_3levels} --> -->

<!-- <!-- ``` --> -->

<!-- *Conclusion*: The test with MMPD selects just one pair, as opposed to the test with maximum. This needs to be checked against what we expect from the test. The harmony pairs which are selected (either through MMPD or maximum) are displayed. -->

@@

# Application

# Conclusion
